Features for detecting threats of violence in YouTube comments

### 2016-03-30 submission.tex
misc smaller adjustments

\title{Features for detecting threats of violence in YouTube comments}
%\title{Detecting threats of violence in YouTube comments}

% Author information can be set in various styles:
% For several authors from the same institution:
\author{Aksel Wester \and Lilja Øvrelid \and Erik Velldal \\
  Department of Informatics \\
  University of Oslo, Norway \\
  ... \\
  \{aksellw, liljao, erikve\}@ifi.uio.no}
% if the names do not fit well on one line use

@@

\begin{abstract}
This article investigates the effect of various types of linguistic features (lexical, morphosyntactic and lexical semantic) for the task of detecting threats of violence in a corpus of YouTube comments. It further examines the use of these features in combination with different representations of linguistic context, and feature backoff. Our results show that combinations of lexical features outperform the use of more complex syntactic and semantic features for this task.
\end{abstract}

\section{Introduction}
\label{sec:intro}
Threats of violence constitute an increasingly common occurrence in online discussions. It disproportionally affects women and minorities, often to the point of effectively eliminating them from taking part in discussions online. Moderators of social networks operate on such a large scale that manually reading all posts is an insurmountable task. Methods for automatically detecting threats could therefore potentially be very helpful, both to moderators of social networks, and to the members of those networks.

@@

There has been quite a bit of work focused on the detection of threats in a data set of Dutch tweets \cite{nellngram2013,nellshallow2013}, which consists of a collection of 5000 threatening tweets collected by a website. In addition, a large number of random tweets were collected for development and testing. %% A set of 2.3 million random tweets was used
%% for development, and a set of 1 million was used for testing.
The system relies on manually constructed recognition patterns, in the form of n-grams 
%(uni-, bi- and trigrams, as well as skip bi- and trigrams)
, but do not go into detail about the methods used to construct these patterns%% , stating that the researchers relied on their
%% (linguistic) intuition as speakers of Dutch \cite{nellngram2013}
. In \newcite{nellshallow2013}, a manually crafted shallow parser is added to the system. This improves results to a precision of 0.39 and a recall of 0.59.
% handcrafted patterns
% rule-based shallow parser
% dubious evaluation

@@

\begin{table}
\begin{center}
\begin{tabular}{lrrr}
  %&   &   &   \\
  \toprule

@@

\end{tabular}
\end{center}
\caption{Number of comments, sentences and users in \ds{}.}
\label{tab:dataset}
\end{table}

@@

In this work we experiment with a wide range of linguistic features in our system and investigate the introduction of linguistic context through a set of feature templates. We also generalize these features through a backoff technique. Throughout, we make use of resources and tools that are freely available and reusable in order to construct our system.

@@

We test three different classifiers in our development testing; a Maximum Entropy (MaxEnt) classifier, a Support Vector Machine (SVM), and a RandomForest classifier (RF).  We approach the task as a binary classification task, and we use scikit-learn \cite{scikit-learn} for our classification.

\paragraph{Tuning}
When tuning each model, we aim to maximize the F-score of the model. In the MaxEnt classifier and SVM we tune the C-value, which for both classifiers is the inverse of regularization strength. When tuning these classifiers, we start with C-values from 1 to 150 in 10-value increments, select the best performing C-value and repeat the process with ever-decreasing increments, with the range of C-values centered on the best performing C-value thus far. After 6 iterations, we terminate the tuning, and select the best performing C-value. When tuning the RandomForest classifier, we perform a grid search over the maximum number of features used when splitting a node in the tree (sqrt($n$), log($n$) or $n$, where $n$ is the number of features in the model), and the number of trees (10, 100 or 1000, depending on the size of the feature set). In the following experiments, we perform tuning on all feature sets.

@@

\item Depdendency triples: \{$<$m, nsubj, i$>$, $<$root, root, m$>$, $<$scared, amod, fucking$>$, ... \}
\end{itemize}
Our lexical features are very specific and require the exact combination of two lexical items in order to apply to a new instance. Following \newcite{Jos:Pen:2009}, we therefore experiment with generalization of features by backoff to a more general category, e.g. from word form to lemma or PoS. A dependency triple over word forms like $<$kill, dobj, them$>$ would thus be generalized to $<$VERB, dobj, them$>$ using \emph{head-backoff}, and $<$kill, dobj, PRON$>$ using \emph{modifier-backoff}. We apply backoff to bigram, trigram and dependency triple features.

@@

\begin{table}
  \begin{center}
    \begin{tabular}{lrrr}
      \toprule

@@

      \bottomrule
      % Data fra firstPass i classify
      % Source/printFirstPassResults.py
    \end{tabular}
  \end{center}
  \caption{Results for baseline system; F-score for bag-of lexical features (word form and lemma), and bigram and trigram templates over word forms . MaxEnt is the MaximumEntropy classifier, SVM is the support vector machine, and RF is the RandomForest classifier.}
  \label{tab:baseline}
\end{table}

@@

    \end{tabular}
  \end{center}
  \caption{F-scores of the bag of words feature set, with different combinations of the other feature sets tested in Table \ref{tab:baseline}.}
  \label{tab:ngrams}
  % tidligere Tabell 4
\end{table}

\begin{table}
  \begin{center}
    \begin{tabular}{lrrrrr}
      \toprule

@@

    \end{tabular}
  \end{center}
  \caption{F-scores of bag of features (BoF) with and without bag of words (BoW). POS\textsubscript{Lex} and Dep\textsubscript{Lex} are lexicalized POS and dep-tags, respectively, where each feature is a tuple consisting of the tag and the word form of the token.}
  \label{tab:bagOfFeatures}
  % tidligere Tabell 3
\end{table}

@@

\begin{table}
  \begin{center}
    \begin{tabular}{lrrrrrr}
      \toprule
      %% BoW, BoF \& 
      %% & Lemma & POS & Dep & Synset & Brown\\
      %% \midrule
      %% Bigram + backoff
      %%  \\ % synset reg, b3200
      %% Trigram + backoff
      %%  \\ % synset reg, b320
      %% $n$-gram + backoff 
      %% \\ % synset reg, b320

      & bigram & trigram & bi+trigram \\
      \midrule
      Lemma backoff & $0.6611$ & $0.6480$ &  $\textbf{0.6649}$ \\ 
      POS backoff & $\textbf{0.6410}$ & $0.6294$ & $0.6320$ \\
      Dep backoff & $0.6208$ & $0.6194$ & $\textbf{0.6220}$ \\
      Synset backoff & $0.6454$ & $0.6448$ &  $\textbf{0.6537}$\\
      Brown backoff & $0.6285$ &  $0.6173$  & $\textbf{0.6335}$\\

@@

    \end{tabular}
  \end{center}
  \caption{Backoff from different combinations of $n$-grams. The models also contain bag of words, and bag of features. Each backoff combination is the one that achieved the best result.}
  \label{tab:ngrambackoff}
\end{table}

@@

\begin{table}
  \begin{center}
    \begin{tabular}{lrr}
      \toprule

      Dependency backoff & Simple & Combined\\
      \midrule
      w/o backoff & $0.6240$ & $\textbf{0.6586}$ \\
      Lemma & $0.6224$ & $0.6507$ \\
      POS & $\textbf{0.6298}$ & $0.6547$ \\ % uPOS, begge to
      Synset & $0.6234$ & $0.6516$  \\ % 1. regular og granparent (de er like), 2. grand parent
      Brown & $0.6299$ & $0.6504$ \\ % 1. Brown 320, 2. Brown 1000

    \end{tabular}
  \end{center}
  \caption{Dependency triples with and without feature backoff. Simple is the feature set containing bag of words and dependency triples. Combined is the feature set containing bag of words, bag of lemmas, bigrams, trigrams and dependency triples. Each row backs off to a different feature type.}
  \label{tab:depbackoff}
\end{table}

@@

\begin{table}
  \begin{center}
    \begin{tabular}{lrrr}
      \toprule

      & Precision & Recall & F-score\\
      \midrule
      Baseline & $0.7325$ & $0.5943$ & $0.6562$ \\
      Combined & $0.7532$ & $\textbf{0.6299}$ & $\textbf{0.6860}$  \\
      Combined+backoff & $\textbf{0.7703}$ & $0.6085$ & $0.6799$ \\

    \end{tabular}
  \end{center}
  \caption{Precision, recall and F-score for the baseline feature set, the best feature set, and the best backoff feature set, when classification is performed on the held-out test set.}
  \label{tab:heldout}
\end{table}

%% \section{Conclusion}
%% \label{sec:disc}

>>>>

\title{Features for detecting threats of violence in YouTube comments}
%\title{Online threat detection}
%\title{Threat detection for online comments}
%\title{Detecting threats of violence in online comments}
%\title{Detecting violent threats online}

% Author information can be set in various styles:
% For several authors from the same institution:
\author{Aksel Wester \and Lilja Øvrelid \and Erik Velldal \\
  Department of Informatics \\
  University of Oslo, Norway \\
  \{aksellw, liljao, erikve\}@ifi.uio.no}
% if the names do not fit well on one line use

@@

\begin{abstract}
This paper investigates the effect of various types of linguistic features (lexical, syntactic and semantic) for the task of detecting threats of violence in a corpus of YouTube comments. It further examines the use of these features in combination with different representations of linguistic context, and feature backoff. Our results show that combinations of lexical features outperform the use of more complex syntactic and semantic features for this task.
\end{abstract}

\section{Introduction}
\label{sec:intro}
Threats of violence constitute an increasingly common occurrence in online discussions. It disproportionally affects women and minorities, often to the point of effectively eliminating them from taking part in discussions online. Moderators of social networks operate on such a large scale that manually reading all posts is an insurmountable task. Methods for automatically detecting threats could therefore potentially be very helpful, both to moderators of social networks, and to their members.

@@

There has been quite a bit of work focused on the detection of threats in a data set of Dutch tweets \cite{nellngram2013,nellshallow2013}, which consists of a collection of 5000 threatening tweets. % collected by a website. 
In addition, a large number of random tweets were collected for development and testing. %% A set of 2.3 million random tweets was used
%% for development, and a set of 1 million was used for testing.
The system relies on manually constructed recognition patterns in the form of n-grams, 
%(uni-, bi- and trigrams, as well as skip bi- and trigrams)
but details about the methods used to construct these patterns are not given.%% , stating that the researchers relied on their
%% (linguistic) intuition as speakers of Dutch \cite{nellngram2013}
In \newcite{nellshallow2013}, a manually crafted shallow parser is added to the system. This improves results to a precision of 0.39 and a recall of 0.59.
% handcrafted patterns
% rule-based shallow parser
% dubious evaluation

@@

\begin{table}
  \begin{smaller}
    \begin{center}
      \begin{tabular}{lrrr}
        % & & & \\
        \toprule

@@

      \end{tabular}
      \caption{Number of comments, sentences and users in \ds{}.}
      \label{tab:dataset}
    \end{center}
  \end{smaller}
\end{table}

@@

In this work we experiment with a wide range of linguistic features in our system and investigate the introduction of linguistic context through a set of feature templates. 
% fixme: det virker litt rart å skille mellom trekk og lingvistisk kontekst? 
We also generalize these features through a backoff technique. Throughout, we make use of resources and tools that are freely available and reusable in order to construct our system.

@@

We test three different classifiers in our development testing; a Maximum Entropy (MaxEnt) classifier, a Support Vector Machine (SVM), and a Random Forest classifier (RF).  We approach the task as a binary classification task, and we use scikit-learn \cite{scikit-learn} for our classification.

\paragraph{Tuning}
When tuning each model, we aim to maximize the F-score of the model. In the MaxEnt classifier and SVM we tune the C-value, which for both classifiers is the inverse of regularization strength. When tuning these classifiers, we start with C-values from 1 to 150 in 10-value increments, select the best performing C-value and repeat the process with ever-decreasing increments, with the range of C-values centered on the best performing C-value thus far. After 6 iterations, we terminate the tuning, and select the best performing C-value. When tuning the Random Forest classifier, we perform a grid search over the maximum number of features used when splitting a node in the tree (sqrt($n$), log($n$) or $n$, where $n$ is the number of features in the model), and the number of trees (10, 100 or 1000, depending on the size of the feature set). In the following experiments, we perform tuning on all feature sets.

@@

\item Depdendency triples: \{$<$m, nsubj, i$>$, $<$root, root, m$>$, $<$scared, amod, fucking$>$, ... \}
\end{itemize}
Our lexical features are very specific and require the exact combination of two lexical items in order to apply to a new instance. Following \newcite{Jos:Pen:2009}, we therefore experiment with generalizing features by `backing off' to a more general category, e.g., from word form to lemma or PoS. For example, a dependency triple over word forms like $<$kill, dobj, them$>$ would thus be generalized to $<$VERB, dobj, them$>$ using \emph{head-backoff}, and $<$kill, dobj, PRON$>$ using \emph{modifier-backoff}. These additional backoff features are included for bigrams and trigrams as well as dependency triples.

@@

\begin{table}
  \begin{smaller}
    \begin{center}
      \begin{tabular}{lrrr}
        \toprule

@@

        \bottomrule
        % Data fra firstPass i classify
        % Source/printFirstPassResults.py
      \end{tabular}
    \end{center}
    \caption{Results for baseline system; F-score for bag-of lexical features
      (word form and lemma), and bigram and trigram templates over word
      forms. From left to right the columns correspond to Maximum Entropy
      classifiers, Support Vector Machines, and Random Forests.}
    \label{tab:baseline}
  \end{smaller}
\end{table}

@@

      \end{tabular}
    \end{center}
    \caption{F-scores of the bag of words feature set, with different
      combinations of the other feature sets tested in Table
      \ref{tab:baseline}.}
    \label{tab:ngrams}
    % tidligere Tabell 4
  \end{smaller}
\end{table}

\begin{table}
  \begin{smaller}    
  \begin{center}
    \begin{tabular}{lrrrrr}
      \toprule

@@

    \end{tabular}
  \end{center}
  \caption{F-scores of bag of features (BoF) with and without bag of words (BoW). POS\textsubscript{Lex} and Dep\textsubscript{Lex} are lexicalized POS and dep-tags, respectively, where each feature is a tuple consisting of the tag and the word form of the token.}
  \label{tab:bagOfFeatures}
  % tidligere Tabell 3
  \end{smaller}
\end{table}

@@

\begin{table}
  \begin{smaller}    
  \begin{center}
    \begin{tabular}{lrrrrrr}
      \toprule
      %% BoW, BoF \& 
      %% & Lemma & POS & Dep & Synset & Brown\\
      %% \midrule
      %% Bigram + backoff
      %%  \\ % synset reg, b3200
      %% Trigram + backoff
      %%  \\ % synset reg, b320
      %% $n$-gram + backoff 
      %% \\ % synset reg, b320
      & bigram & trigram & bi+trigram \\
      \midrule
      Lemma backoff & $0.6611$ & $0.6480$ &  $\textbf{0.6649}$ \\ 
      POS backoff & $\textbf{0.6410}$ & $0.6294$ & $0.6320$ \\
      Dep backoff & $0.6208$ & $0.6194$ & $\textbf{0.6220}$ \\
      Synset backoff & $0.6454$ & $0.6448$ &  $\textbf{0.6537}$\\
      Brown backoff & $0.6285$ &  $0.6173$  & $\textbf{0.6335}$\\

@@

    \end{tabular}
  \end{center}
  \caption{Backoff from different combinations of $n$-grams. The models also contain bag of words, and bag of features. Each backoff combination is the one that achieved the best result.}
  \label{tab:ngrambackoff}
  \end{smaller}
\end{table}

@@

\begin{table}
  \begin{smaller}    
  \begin{center}
    \begin{tabular}{lrr}
      \toprule           
      Dependency backoff & Simple & Combined\\
      \midrule
      w/o backoff & $0.6240$ & $\textbf{0.6586}$ \\
      Lemma & $0.6224$ & $0.6507$ \\
      POS & $\textbf{0.6298}$ & $0.6547$ \\ % uPOS, begge to
      Synset & $0.6234$ & $0.6516$  \\ % 1. regular og granparent (de er like), 2. grand parent
      Brown & $0.6299$ & $0.6504$ \\ % 1. Brown 320, 2. Brown 1000                 
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{Dependency triples with and without feature backoff. Simple is the feature set containing bag of words and dependency triples. Combined is the feature set containing bag of words, bag of lemmas, bigrams, trigrams and dependency triples. Each row backs off to a different feature type.}
  \label{tab:depbackoff}
  \end{smaller}
\end{table}

@@

\begin{table}
\begin{smaller}
  \begin{center}
    \begin{tabular}{lrrr}
      \toprule      
      & Precision & Recall & F-score\\
      \midrule
      Baseline & $0.7325$ & $0.5943$ & $0.6562$ \\
      Combined & $0.7532$ & $\textbf{0.6299}$ & $\textbf{0.6860}$  \\
      Combined+backoff & $\textbf{0.7703}$ & $0.6085$ & $0.6799$ \\            
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{Precision, recall and F-score for the baseline feature set, the best feature set, and the best backoff feature set, when classification is performed on the held-out test set.}
  \label{tab:heldout}
\end{smaller}
\end{table}

\section{Conclusion}
\label{sec:conclusion}
